---
title: "Social Survey"
author: "Vanessa Michel"
date: "November 13, 2017"
output:
  html_document: default
---

```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
source("Import_GSS_2014_haven.R")
```
<!-- We recommend you use git branches -- create a separate branch for each major section i.e. Introduction, Methods, Results, Inferential Results, and Discussion and Conclusions - leave the YAML block untouched until the final merge where you put in your own name !-->

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->   

####This study deals with questions about possible relationships between gun ownership and race, sex, political views, opinions on capitol punishment and belief in god. The data is from the 2014 general social survey.

We will use the data from this survey to answer the following research question:  
Is there a relationship between gun ownership and a person's race, sex, political views, opinion on capitol pushment and belief in god?

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->
We will be using the folloiwng methods: bar charts, two-way tables and a chi-squared test. We will use these to analyze the following factor variables: whether or not someone owns a gun, a person's race, sex, political views, opinion on capitol punishment and beilef in god.

#Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
##### **Gun ownership vs. Race Barchart**
```{r}
race_table=xtabs(~race + owngun, data=n2014,exclude = "NA")
barchartGC(race_table,type="percent")
```

After viewing the data graphically, it appears that there are more White gun owners than there are Black or "Other" gun owners. 

##### **Gun ownership vs. Sex Barchart**
```{r}
sex_table=xtabs(~sex + owngun, data=n2014,exclude = "NA")
barchartGC(sex_table,type="percent")
```

It appears that there is a relationship between sex and gun ownership. Males seem to be more likely to own a gun than females are. 

##### **Gun onership vs. Political Views Barchart**
```{r}
polviews_table=xtabs(~polviews + owngun, data=n2014,exclude = "NA")
barchartGC(polviews_table,type="percent")
```

It appears that there is a relationship seen between a person's political view and whether or not one own's a gun. Conservatives are more likely to own a gun. Liberals are much less likely to own a gun. Moderate people are in the middle, where about half of them are gun owners. 

##### **Gun ownership vs. Opinion on Capitol Punishment Barchart**
```{r}
cappun_table=xtabs(~cappun + owngun, data=n2014,exclude = "NA")
barchartGC(cappun_table,type="percent")
```

There appears to be a strong relationship between a person's view on capitol punishment and gun ownership. Those who favor capitol punishment are much more likely to own a gun than those who oppose it. 

##### **Gun ownership vs. Belief in god Barchart**
```{r}
god_table=xtabs(~god + owngun, data=n2014,exclude = "NA")
barchartGC(god_table,type="percent")
```

There appears to be no relationship between whether or not a person believes in god and whether or not they own a gun. In each category it seems about half the people belonging in that group own a gun. The other half do not. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

To view the results of our data numerically, we will use a two-way table with row percents. 

##### **Gun ownership vs. Race Two-way Table:**
```{r}
rowPerc(race_table)
```
Of those in the **WHITE** group, 39.01% own a gun, 60.99% do not.  
Of those in the ***BLACK** group, 16.87% own a gun, 83.13% do not.   
Of those in the **OTHER** group, 13.93% own a gun, 86.07% do not.  
After viewing the data numerically, it appears that a higher percentage of **White** tend to own guns. For **Black** and **Other** more tend to not own a gun vs. those that do.

##### **Gun ownership vs. Sex Two-way Table:**
```{r}
rowPerc(sex_table)
```
Of those in the **MALE** group, 39.76% own a gun, 60.24% do not.  
Of those in the **FEMALE** group, 27.32% own a gun, 72.68% do not.  
After viewing the data numerically, it appears that a higher percentage of **Male** participants own a gun as opposed to the **Female** participants.

##### **Gun ownership vs. Political View Two-way Table:**
```{r}
rowPerc(polviews_table)
```


##### **Gun ownership vs. Opinion on Capitol Punishment Two-way Table:**
```{r}
rowPerc(cappun_table)
```


##### **Gun ownership vs. Belief in god Two-way Table:**
```{r}
rowPerc(god_table)
```

# Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
